\documentclass[letterpaper, 10pt]{article}
%\usepackage{fullpage} % changes the margin
\usepackage{geometry}
 \geometry{
 letterpaper,
 total={170mm,244mm},
 left=22mm,
 top=14mm,
 }
 \usepackage[UKenglish]{isodate}% http://ctan.org/pkg/isodate for date format
\usepackage{wrapfig}
\usepackage[font={small,sf}]{caption}
\usepackage{float}%force tables/figs into certain placement
\usepackage{graphicx}%for figures
\usepackage[leftcaption]{sidecap}%for figure captions
\usepackage{hyperref}%for hyperlinks
\usepackage[font=small,labelfont=bf]{caption}%for captions
\usepackage{natbib}	%for bibliography


\def\labelitemi{--}

\title{\textbf{The Insect Collectors' Code}\footnote{This oath was modified from the modern version of the Hippocratic oath, written in 1964 by Louis Lasagna and available at: \url{http://www.aapsonline.org/ethics/oaths.htm\#lasagna} (accessed 26 April 2017)}}
\author{Carolyn Trietsch \& Andrew R. Deans\footnote{contact:  carolyntrietsch@gmail.com (CT), adeans@psu.edu (ARD)}\\Frost Entomological Museum, Pennsylvania State University, University Park, PA  USA}

\begin{document}
\cleanlookdateon %removed ordinal date
\maketitle

\noindent{}I strive to fulfill, to the best of my ability, the following ideals:

\begin{enumerate}
\item I will respect the hard-won scientific gains of those entomologists in whose steps I walk and gladly share my scientific gains and knowledge with those who are to follow.
\item I will aid in the dissemination of scientific knowledge, both to those who study insects and those who do not.
\item I will not discriminate against others, and I will strive to create a safe working environment, whether in the field, the classroom, or the lab.
\item I will treat insects humanely. As a collector, it is within my power to take insect life; I will not take insects that will not be deposited in a natural history collection or otherwise made available for research and education. While bycatch is often unavoidable, I will, to the best of my ability, attempt to reduce the unnecessary loss of insect life and find use for these specimens.
\item I will consider the ecological impact of removing insects and their products (galls, nests, \textit{etc}.) from the environment when collecting, whether the species are protected by law, known to be declining, or are considered to be of least concern. I will strive to avoid or minimize disturbance to the environment while collecting.
\item I will secure appropriate permits prior to collecting insects, and I will honor and uphold the provisions stated by each permit. I will keep copies of all permits on my person while collecting and furnish them to authorized agents upon request. I will save all permits associated with specimens as proof that they were collected legally.
\item I will keep detailed field notes of my collecting activities and will make these available to the greater scientific community.
\item I will prepare and label specimens according to standards established by professional entomologists who work with collections.
\item I will properly store all specimens under my care, and I will not allow specimens to become damaged or degraded through neglect.
\item I will properly use and dispose of preservatives, killing agents, and other chemicals associated with specimen collection and preparation. I will never use these chemicals to harm myself or others.
\item I will make arrangements for any personal specimens or collections in my possession to be deposited in a museum in the event of my untimely death.
\item I will not create false data.
\end{enumerate}
 
\noindent{}May I always act so as to preserve the finest traditions of natural history, and so long as I uphold these traditions and the stated ideals, may I long experience the joy of my contributions to the furthering of scientific knowledge. 
\vfill

\end{document}
